---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

## Libraries 
```{r}
library("dplyr")
```




## Loading and preprocessing the data
### Unzip Data File
```{r}
if(!file.exists("activity.csv"))
{
  unzip("activity.zip")
}
```
### Load data
```{r}
dataActivity <- read.csv("activity.csv")
```


### Some summary statistics
```{r}
summary(dataActivity)
```

How many days exists in data set?

```{r}
unique(dataActivity$date)
```




## What is mean total number of steps taken per day?


```{r}
dataActivityNARemoved <- dataActivity[complete.cases(dataActivity),]

totalsGroupedByDate <- dataActivityNARemoved %>%
  group_by(date) %>%
  select(steps, date) %>%
  summarise(
    sumStep = sum(steps, na.rm = TRUE)
  ) 
breakSize <-dim(totalsGroupedByDate)[1]
hist(totalsGroupedByDate$sumStep,breaks=breakSize,main="Total Steps taken each day",ylab="totalSteps",xlab="date")
```

### mean 
```{r}
mean(totalsGroupedByDate$sumStep)
```

### median 

```{r}
median(totalsGroupedByDate$sumStep)
```



## What is the average daily activity pattern?
```{r}
averageGroupedByInterval <- dataActivity %>%
  group_by(interval) %>%
  select(steps, interval) %>%
  summarise(
    averageStep = mean(steps, na.rm = TRUE)
  ) 

plot(averageGroupedByInterval$interval,averageGroupedByInterval$averageStep,type="l",xlab="intervals",ylab="average steps for every day",main="average number of steps taken each 5 minute interval")

```
### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxSteps <- averageGroupedByInterval  %>%
  arrange( desc(averageStep)) 

maxSteps[ 1,]
```

## Imputing missing values

### total number of missing values in the dataset 

```{r}
howManyNAinSteps <- sum(!complete.cases(dataActivity[,1]))
howManyNAinSteps
howManyNAinDate <- sum(!complete.cases(dataActivity[,2]))
howManyNAinDate
howManyNAinInterval <- sum(!complete.cases(dataActivity[,3]))
howManyNAinInterval
```

Only missing values are in steps, no missing values for other features.

Devise a strategy for filling in all of the missing values in the dataset.
We use the mean for that 5-minute interval.

```{r}
dataActivityMissingValuesImputed <- dataActivity
dataActivityMissingValuesImputed[!complete.cases(dataActivity[,1]),1] <- averageGroupedByInterval[averageGroupedByInterval[,1] == dataActivity[!complete.cases(dataActivity[,1]),3],2]$averageStep

```

Calculate Histogram with NAs imputed

```{r}

totalsGroupedByDateNAImputed <- dataActivityMissingValuesImputed %>%
  group_by(date) %>%
  select(steps, date) %>%
  summarise(
    sumStep = sum(steps, na.rm = TRUE)
  ) 
breakSize <-dim(totalsGroupedByDateNAImputed)[1]
hist(totalsGroupedByDateNAImputed$sumStep,breaks=breakSize,main="Total Steps taken each day",ylab="totalSteps",xlab="date")

```

### mean 
```{r}
mean(totalsGroupedByDateNAImputed$sumStep)
```

### median 

```{r}
median(totalsGroupedByDateNAImputed$sumStep)
```

### Do these values differ from the estimates from the first part of the assignment? 

```{r}
mean(totalsGroupedByDateNAImputed$sumStep) - mean(totalsGroupedByDate$sumStep)

median(totalsGroupedByDateNAImputed$sumStep) - median(totalsGroupedByDate$sumStep)

```
Mean does not change due to our strategy to replace missing values with mean.
But median changes.

### What is the impact of imputing missing data on the estimates of the total daily number of steps?

Total daily number of steps increases since we replaced NA values with mean.






## Are there differences in activity patterns between weekdays and weekends?
